Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-22 robdml.tex
tex

%%%%%%%%% ABSTRACT
\begin{abstract}
	% background
	Owing to the security implications of adversarial vulnerability, the
	adversarial robustness of deep metric learning models have to be improved
	via defense.
	% insight
	In order to avoid model collapse due to excessively hard examples, the
	existing methods refrain from using the typical min-max adversarial
	training paradigm, but instead learn from a weak adversary.
	% our starting point
	Conversely, we argue that a model can learn from a strong adversary without
	collapse, where the key is triplet hardness.
	% HM framework
	To this end, Hardness Manipulation is proposed to adversarially perturb the
	training triplet to increase its hardness into a specified level, by
	referencing another benign triplet or a pseudo-hardness function.
	% GA Adversary
	Based on this, Gradual Adversary is proposed as a family of pseudo-hardness
	functions to overcome problems of referencing other benign triplets, and
	achieve a better balance in the learning objective.
	% Intra-Class Structure
	Besides, a penalty on Intra-Class Structure among benign and adversarial
	examples can further improve model robustness, which is neglected by
	previous methods.
	% Experiment and Conclusion
	Comprehensive experimental results on three commonly used deep metric
	learning datasets suggest that the proposed method overwhelmingly
	outperforms the state-of-the-art defense methods in terms of robustness,
	training cost, as well as recall performance on benign examples, while
	remains in a simple form.
%
\end{abstract}

@@

Given a set of data points, a metric gives a distance value between each pair
of them.
%
Deep Metric Learning (DML) aims to learn such a metric between two inputs (\eg,
images) leveraging the representational power of deep neural networks.
%
As an extensively studied task~\cite{revisiting}, DML has a wide range of
applications such as image retrieval~\cite{imagesim2} and face
recognition~\cite{facenet,domainface}, as well as a wide impact in
the core idea of other areas such as self-supervised learning~\cite{dmlreality}.

Despite the improvements of this field thanks to the advancement of deep neural
networks, recent research works suggest that DML models are vulnerable to
adversarial attacks, where an imperceptible perturbation could incur unexpected
or covertly manipulated results~\cite{advrank,advorder}.
%
Such vulnerability raises safety, security, and fairness concerns in the
applications of DML.
%
For example, impersonation or recognition evation are possible for DML-based
face-identification system.
%
To counter the attacks (\ie, mitigating the vulnerability), it is important to
design defense methods to improve the adversarial robustness of DML models.

@@

% existing methods & problem
Several defense methods for DML are proposed in the
literature~\cite{advrank,robrank}, inspired by the Madry's adversarial training
method~\cite{madry}, as it is kown as one of the most effective defense methods
for deep neural network classifiers.
%
However, it has been noted that
%
(1) the robustness level achieved by the existing methods is still insufficient
to counter the attacks;
%
(2) the direct adaptation of Madry's min-max adversarial training~\cite{madry} will easily
lead to model collapse due to producing very hard sample triplets;
%
(3) the adversarial training procedure is very time-consuming compared to
the training process of a regular DML model, but existing defense methods
are incompatible to acceleration methods like Free Adversarial Training~\cite{freeat}.
\oo{rob, time, r1}.

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
	% background
	Owing to the security implications of adversarial vulnerability, the
	adversarial robustness of deep metric learning models have to be improved
	via defense.
	% insight
	In order to avoid model collapse due to excessively hard examples, the
	existing methods refrain from using the min-max adversarial
	training, but instead learns from a weak adversary inefficiently.
	% our starting point
	Conversely, we argue that a model can learn from a strong adversary without
	collapse, where the key is triplet hardness.
	% HM framework
	To this end, Hardness Manipulation is proposed to adversarially perturb the
	training triplet to increase its hardness into a specified level, by
	referencing another benign triplet or a pseudo-hardness function.
	% GA Adversary
	Based on this, Gradual Adversary is proposed as a family of pseudo-hardness
	functions to overcome problems of referencing other benign triplets, and
	achieve a better balance in the learning objective.
	% Intra-Class Structure
	Besides, a penalty on Intra-Class Structure among benign and adversarial
	examples can further improve model robustness, which is neglected by
	previous methods.
	% Experiment and Conclusion
	Comprehensive experimental results on three commonly used deep metric
	learning datasets suggest that the proposed method overwhelmingly
	outperforms the state-of-the-art defense methods in terms of robustness,
	training cost, as well as recall performance on benign examples, while
	remains in a simple form.
%
\end{abstract}

@@

Given a set of data points, a metric gives a distance value between each pair
of them.
%
Deep Metric Learning (DML) aims to learn such a metric between two inputs (\eg,
images) leveraging the representational power of deep neural networks.
%
As an extensively studied task~\cite{revisiting}, DML has a wide range of
applications such as image retrieval~\cite{imagesim2} and face
recognition~\cite{facenet,domainface}, as well as a wide influence on
the core idea of other areas such as self-supervised learning~\cite{dmlreality}.

Despite the advancements in this field thanks to deep neural
networks, recent studies find DML models vulnerable to
adversarial attacks, where an imperceptible perturbation can incur unexpected
retrieval, or covertly manipulate the results~\cite{advrank,advorder}.
%
Such vulnerability raises security, safety, and fairness concerns in the
DML applications.
%
For example, impersonation or recognition evation are possible for a
vulnerable DML-based face-identification system.
%
To counter the attacks (\ie, mitigating the vulnerability), the
adversarial robustness of DML models has to be improved via defense.

@@

% existing methods & problem
Several adversarial training-based defense methods for DML are proposed in the
literature~\cite{advrank,robrank}, which are inspired by the Madry's min-max
adversarial training~\cite{madry} because it is one of the most effective
defense methods for deep neural network classifiers.
%
However, it is widely known that such methods always suffer from greatly increased
training cost and a notable penalty in the model's retrieval performance on
benign (\ie, unperturbed) examples.
%
Meanwhile, the level of robustness that the previous methods have achieved
is still insufficient.
%
(1) the robustness level achieved by the existing methods is still insufficient
to counter the attacks;
%
(2) the direct adaptation of Madry's min-max adversarial training~\cite{madry}
will easily lead to model collapse due to producing very hard sample triplets;
%
(3) the adversarial training procedure is very time-consuming compared to the
training process of a regular DML model, but existing defense methods are
incompatible to acceleration methods like Free Adversarial
Training~\cite{freeat}.  \oo{rob, time, r1}.
